---
title: "COTS_ProjectOrganization"
author: "Kevin Shoemaker, Sam Matthews, Camille Mellin, Damien Fordham, ..."
date: "April 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This document is intended to describe the overall project organization for constructing an R-based population model for COTS in the Great Barrier Reef. 

# Dropbox: shared project resources
All shared project resources are stored in [Dropbox](www.Dropbox.com). There are several key folders for storing different types of resources that are read into the R-based model or produced as outputs from the R-based model. 

## Data
This folder is a repository for storing CSV files for reading into the R COTS modeling framework. Key variables for cross-referencing data files include:

* PID (population ID): a unique ID for each 1 km grid cell containing some coral reef.
* RID (reef ID): a unique ID for each separate reef in the Great Barrier Reef.

## Spatial Layers
This folder contains all shared spatial resources. Raster layers are stored as .ASC grids, and vector layers are stored as ESRI Shapefiles. Note that raster layers are stored directly in the "Spatial layers" folder, but shapefiles (which comprise multiple component files) are stored in subfolders. Key spatial layers include:

* GBR_reefs: Shapefile indicating the locations and IDs of all coral reefs in the GBR
* templateRaster.asc: ascii raster file containing no data but indicating the proper grid parameters for the study area (extent and resolution)
* reefPercentRaster.asc: ascii raster file indicating, for each pixel, the percent of coral reef coverage

## Figures
This folder contains figures and plots for interpreting results. The subfolder "RawFigures" contains figures generated directly from R. Final figures may be generated using a vector graphics software such as Inkscape. 

## R_Workspaces
This folder contains time-stamped R workspaces (.RData files) to enable users to pick up where they left off.  



# GIThub: source code
We are using [GIThub](github.com) for version control for all source code. The desktop software can be downloaded [here](https://desktop.github.com/)


NOTE: before any project code can be run, the following r script must be run, to ensure that R knows where to find the resource files and source code files, and to load all functions and libraries associated with this project. 

```{r eval=FALSE}
#######################
#   CLEAR THE WORKSPACE 
#######################

rm(list=ls())

#######################
#   SET USER 
#######################

USER = "KEVIN"

#########################
# SET PROJECT DIRECTORIES (this should be the only place where local directories should be referenced)
#########################

if(USER=="KEVIN") BASE_DIRECTORY <- "C:\\Users\\Kevin\\Dropbox\\CoTS_Model"             # NOTE: this should link to the Dropbox folder with shared project resources	                                                                        
if(USER=="KEVIN") CODE_DIRECTORY <- "C:\\Users\\Kevin\\GIT\\COTS_popmodel"              # NOTE: code directory should be your local copy of the GitHub repository

if(USER=="SAM") BASE_DIRECTORY ##<- ****FILL IN*****
if(USER=="SAM") CODE_DIRECTORY ##<- ****FILL IN*****

SPATIALDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Spatial Layers",sep="")                          # directory for storing relevant spatial data (ASC, SHP files)
if(is.na(file.info(SPATIALDATA_DIRECTORY)[1,"isdir"])) dir.create(SPATIALDATA_DIRECTORY)

DATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Data",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(DATA_DIRECTORY)[1,"isdir"])) dir.create(DATA_DIRECTORY)

ENVDATA_DIRECTORY <- paste(DATA_DIRECTORY,"\\Environmental",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(ENVDATA_DIRECTORY)[1,"isdir"])) dir.create(ENVDATA_DIRECTORY)

FIGURES_DIRECTORY <- paste(BASE_DIRECTORY,"\\Figures\\RawFigures",sep="")               # directory for storing raw figures generated from R
if(is.na(file.info(FIGURES_DIRECTORY)[1,"isdir"])) dir.create(FIGURES_DIRECTORY)

RESULTS_DIRECTORY <- paste(BASE_DIRECTORY,"\\results",sep="")                           # directory for storing relevant results
if(is.na(file.info(RESULTS_DIRECTORY)[1,"isdir"])) dir.create(RESULTS_DIRECTORY)

RDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\R_Workspaces",sep="")                        # directory for storing .RData files (R workspaces and data objects)
if(is.na(file.info(RDATA_DIRECTORY)[1,"isdir"])) dir.create(RDATA_DIRECTORY)

####################
#  LOAD FUNCTIONS AND SCRIPTS FROM SOURCE CODE (your local GitHub repository)
#################### 

setwd(CODE_DIRECTORY)
source("COTSModel_Utilityfunctions.R")   # load utility functions, e.g., for loading packages etc. 
#source("COTSModel_COTSfunctions.R")      # load functions for implementing COTS demography and dispersal
#source("COTSModel_Coralfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
source("COTSModel_GISfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
source("COTSModel_Initialization.R")     # initialize the workspace
```

