---
title: "COTS Population Model, Detailed Description"
author: "Sam Matthew, Camille Mellin, Damien Fordham, Kevin Shoemaker, ..."
date: "April 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)    # set up knitr

#######################
#   CLEAR THE WORKSPACE 
#######################

rm(list=ls())

#########################
# SET PROJECT DIRECTORIES (this should be the only place where local directories should be referenced)
#########################

USER = "KEVIN"

if(USER=="KEVIN") BASE_DIRECTORY <- "C:\\Users\\Kevin\\Dropbox\\CoTS_Model"             # NOTE: this should link to the Dropbox folder with shared project resources	                                                                        
if(USER=="KEVIN") CODE_DIRECTORY <- "C:\\Users\\Kevin\\GIT\\COTS_popmodel"              # NOTE: code directory should be your local copy of the GitHub repository

if(USER=="SAM") BASE_DIRECTORY ##<- ****FILL IN*****
if(USER=="SAM") CODE_DIRECTORY ##<- ****FILL IN*****

SPATIALDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Spatial Layers",sep="")                          # directory for storing relevant spatial data (ASC, SHP files)
if(is.na(file.info(SPATIALDATA_DIRECTORY)[1,"isdir"])) dir.create(SPATIALDATA_DIRECTORY)

DATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\Data",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(DATA_DIRECTORY)[1,"isdir"])) dir.create(DATA_DIRECTORY)

ENVDATA_DIRECTORY <- paste(DATA_DIRECTORY,"\\Environmental",sep="")                                 # directory for storing data (CSV files)
if(is.na(file.info(ENVDATA_DIRECTORY)[1,"isdir"])) dir.create(ENVDATA_DIRECTORY)

FIGURES_DIRECTORY <- paste(BASE_DIRECTORY,"\\Figures\\RawFigures",sep="")               # directory for storing raw figures generated from R
if(is.na(file.info(FIGURES_DIRECTORY)[1,"isdir"])) dir.create(FIGURES_DIRECTORY)

RESULTS_DIRECTORY <- paste(BASE_DIRECTORY,"\\results",sep="")                           # directory for storing relevant results
if(is.na(file.info(RESULTS_DIRECTORY)[1,"isdir"])) dir.create(RESULTS_DIRECTORY)

RDATA_DIRECTORY <- paste(BASE_DIRECTORY,"\\R_Workspaces",sep="")                        # directory for storing .RData files (R workspaces and data objects)
if(is.na(file.info(RDATA_DIRECTORY)[1,"isdir"])) dir.create(RDATA_DIRECTORY)

####################
#  LOAD FUNCTIONS FROM SOURCE CODE (use your local GitHub repository)
#################### 

setwd(CODE_DIRECTORY)
source("COTSModel_Utilityfunctions.R")   # load utility functions, e.g., for loading packages etc. 
#source("COTSModel_COTSfunctions.R")      # load functions for implementing COTS demography and dispersal
#source("COTSModel_Coralfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
#source("COTSModel_GISfunctions.R")     # load functions for implemeting coral growth/recovery and dispersal
source("COTSModel_Initialization.R")     # initialize the workspace

```

# COTS Population Model: objectives

This population model is constructed in R, using the RStudio development environment. 
Yadda yadda **yadda** *yadda* ~~yadda~~ [yadda](www.rstudio.com)


#  TODO LIST

* how to initialize the COTS population??? [KTS developed a crude method using reef-level densities...]
* determine how exactly the coral/cots interaction will work? 
    + is there a maximal live coral cover for each grid cell? What determines this maximum
    + We have cover for two species of coral- how does this enter and influence the model?
    + Is the SDM just used to specify initial live coral cover? Or, does this SDM affect the
  model in a more long-lasting way?


# DATA DESCRIPTION
The population model uses data of three major types: 

1. GIS Rasters: 0.01 degrees (ca. 1 km) resolution. Approx 1.3 million pixels. We chose to use a 1 km resolution to represent geographic variation within the Great Barrier Reef.
We chose this resolution because it was the coarsest resolution that could successfully capture
the vast majority of the small reefs in the landscape.  
2. Population-level data: XYZ dataframe containing unique information (Z1, Z2, ...) about
each 1 km^2^ reef segment, or "population" (ca. 17k populations).  
3. Reef-level data: XYZ dataframe containing unique information (z1, Z2, ...) about 
each reef (variable sizes) (ca. 2k reefs)

### Initial coral cover (GIS, population-level)
This layer represents the was derived from several sources. 




# RMARKDOWN EXAMPLES- FOR REFERENCE
inline equation: $A = \pi*r^{2}$

Here is an example of embedded R code:

```{r}
#a<-2+2
a

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
